---
title: "2012 Wave of HRS 1-sample Mendelian Randomization Assumption Testing"
author: "Kyle Abraham Campbell"
date: '2022-09-17'
output: html_document
---

```{r setup, include=FALSE}
library(boot)
library(formulaic)
library(ivtools)
library(knitr)
library(ggpubr)
library(gtsummary) 
library(here)
library(parameters)
library(MendelianRandomization)
library(readxl)
library(skimr)
library(tidymodels)
library(tidyverse)
```

as.data.frame is ESSENTIAL for ivglm to accept these as valid data inputs
```{r}
#dat <- readRDS(here("data", "analytic", "2022-12-09.current.consumption.gscan.dpw.1smr.xs.rda"))
dat <- readRDS(here("data", "analytic", "2023-02-17.gscan.dpw.1smr.xs.rda"))

# Previous 99 SNP with with chr:pos:ref:alt naming convention
#dat <- readRDS(here("data", "analytic", "2022-09-18.gscan.dpw.1smr.xs.rda")) %>% as_tibble(.name_repair = "universal") #%>% mutate(DEGREE = factor(DEGREE))

#dat.ancestry <- readRDS(here("data", "analytic", "genetics_confirmed_2012_2022-09-18.rda")) %>% filter(RACE %in% c("Non-Hispanic Black, African ancestry", "Non-Hispanic White, European ancestry"))
```

Get SNP IDs

Existing rsXXXX columns in HRS data; need to drop
hrs$rs429358
hrs$rs7412
```{r}
variants <-
  dat %>%
  # This approach assumes the SNP name still includes :'s
  #dplyr::select(matches(":\\D:\\D")) %>%
  # This approach uses updated repaired names
  #dplyr::select(matches("\\.\\D\\.\\D")) %>%
  # This approach uses the updated rsID
  dplyr::select(matches("^rs[0-9]+")) %>%
  colnames()
variants <- variants[!variants %in% c("rs429358", "rs7412")]

univariates = c("HHIDPN", "NAGE", "GENDER", "RACE", "DEGREE", "AD12", "drinkpweek", "logdrinkpweek",
                "R11DRINK", "R11DRINKD", "R11DRINKN", "APOE012",
                "R11SMOKEV", "NMARST", "R11CESD", "R11CONDE", "R11SAYRET",
                "PGS4_DPW_GSCAN19", "pgs.norm", "AD12.CIND.outcome", "AD12.Dementia.outcome", "AD12.binary") 
combined.pgs.covariates <- c("AncestryPC_1_5A", "AncestryPC_1_5B", "AncestryPC_1_5C", "AncestryPC_1_5D", "AncestryPC_1_5E")

model.vars <- c(univariates, combined.pgs.covariates, variants) %>% unique()
genetics <- c("PGS4_DPW_GSCAN19", "pgs.norm", variants)
```

The different kinds of models we build beyond naive and combined.pgs.covariates Ancestry-adjusted 5 PCs
```{r}
demographics <- c("NAGE", "GENDER", "DEGREE")
additional_covars <- c( "R11SMOKEV", "NMARST", "R11CESD", "R11CONDE", "R11SAYRET")
```

```{r}
dat <-
  dat %>%
  # Subset to only the relevant columns
  #dplyr::select(all_of(model.vars)) %>%
  mutate(AD12.binary = factor(AD12.binary))
```

Option to convert to long format with SNP as a single column
```{r, eval = F, include = F}
dat.long <-
  dat %>%
  pivot_longer(all_of(c(variants, "PGS4_DPW_GSCAN19", "pgs.norm")), names_to = "SNP", values_to = "Dosage")
```

```{r}
dat.ea <- dat %>% filter(RACE == "Non-Hispanic White, European ancestry") %>% as.data.frame()
dat.aa <- dat %>% filter(RACE == "Non-Hispanic Black, African ancestry") %>% as.data.frame()

#dat.all.ea <- dat.all %>% filter(RACE == "Non-Hispanic White, European ancestry") %>% as.data.frame()
#dat.all.aa <- dat.all %>% filter(RACE == "Non-Hispanic Black, African ancestry") %>% as.data.frame()
```

# Relevance Assumption 

## Set models and helper functions
Source code for relevance model fitting
```{r}
source(here("scripts", "fit_relevance_models.R"))
```

Set linear regression tidy model
```{r}
lm_mod <-
  linear_reg()
```

Set logistic regression tidy model
```{r}
lr_mod <- 
  logistic_reg() %>% 
  set_engine("glm")
```

## European ancestries
### Drinks per week exposure (log-transformed)
Fit all relevance models for naive, ancestryPCs, demographic, and full models
```{r}
res_relevance_ea <- fit_relevance_models(instruments = genetics, exposure = "logdrinkpweek", tidy_model = lm_mod, data = dat.ea)
```

PGS results
```{r}
print(res_relevance_ea %>% filter(Instrument == "PGS4_DPW_GSCAN19"))
print(res_relevance_ea %>% filter(Instrument == "pgs.norm"))
```


Display and plot results
```{r}
ggplot(data = res_relevance_ea %>% filter(str_detect(Instrument, "^rs[0-9]")),
       mapping = aes(x = graph_F, color = Model, fill = Model)) +
  geom_histogram() +
  facet_wrap(facets = vars(Model)) +
  ggtitle(ggtitle(paste0("European ancestries histogram of instrument F-statistics, n = ", dim(dat.ea)[1])))
```

Report descriptives
```{r}
res_relevance_ea %>% filter(str_detect(Instrument, "^rs[0-9]")) %>% filter(Model == "Full") %>% pull(graph_F) %>% mean
res_relevance_ea %>% filter(str_detect(Instrument, "^rs[0-9]")) %>% filter(Model == "Full") %>% pull(graph_F) %>% sd

res_relevance_ea %>% filter(str_detect(Instrument, "^rs[0-9]")) %>% filter(Model == "Demographics") %>% pull(graph_F) %>% mean
res_relevance_ea %>% filter(str_detect(Instrument, "^rs[0-9]")) %>% filter(Model == "Demographics") %>% pull(graph_F) %>% sd
```


```{r}
ggplot(data = dat.ea, mapping = aes(x = PGS4_DPW_GSCAN19, y = logdrinkpweek)) + geom_point() + geom_smooth(method='lm', formula= y~x) +
  stat_cor() +
  ggtitle(paste0("European ancestries regress DPW on PGS4_DPW_GSCAN19, n = ", dim(dat.ea)[1]))

ggplot(data = dat.ea, mapping = aes(x = pgs.norm, y = logdrinkpweek)) + geom_point() + geom_smooth(method='lm', formula= y~x) +
  stat_cor() +
  ggtitle(paste0("European ancestries regress log-DPW on pgs.norm, n = ", dim(dat.ea)[1]))
```

## African ancestries population - relevance
### Drinks per week exposure

Fit all relevance models for naive, ancestryPCs, demographic, and full models
```{r}
res_relevance_aa <- fit_relevance_models(instruments = genetics, exposure = "logdrinkpweek", tidy_model = lm_mod, data = dat.aa)
```

Report descriptives
```{r}
res_relevance_aa %>% filter(str_detect(Instrument, "^rs[0-9]")) %>% filter(Model == "Demographics") %>% pull(graph_F) %>% mean
res_relevance_aa %>% filter(str_detect(Instrument, "^rs[0-9]")) %>% filter(Model == "Demographics") %>% pull(graph_F) %>% sd

res_relevance_aa %>% filter(str_detect(Instrument, "^rs[0-9]")) %>% filter(Model == "Unadjusted") %>% pull(graph_F) %>% mean
res_relevance_aa %>% filter(str_detect(Instrument, "^rs[0-9]")) %>% filter(Model == "Unadjusted") %>% pull(graph_F) %>% sd
```

Display and plot results
```{r}
ggplot(data = res_relevance_aa,
       mapping = aes(x = graph_F, color = Model, fill = Model)) +
  geom_histogram() +
  facet_wrap(facets = vars(Model)) +
  ggtitle(ggtitle(paste0("African ancestries histogram of instrument F-statistics, n = ", dim(dat.aa)[1])))

print(res_relevance_aa %>% filter(Instrument == "PGS4_DPW_GSCAN19"))

print(res_relevance_aa %>% filter(Instrument == "pgs.norm"))

ggplot(data = dat.aa, mapping = aes(x = PGS4_DPW_GSCAN19, y = drinkpweek)) + geom_point() + geom_smooth(method='lm', formula= y~x) +
  ggtitle(paste0("African ancestries regress DPW on PGS4, n = ", dim(dat.aa)[1]))
```

## Relevance figure

```{r}
geom.point.alpha = 0.5

relevance.dat.ea <- res_relevance_ea %>% mutate(Ancestry = "European ancestry")
relevance.dat.aa <- res_relevance_aa %>% mutate(Ancestry = "African ancestry")

relevance.dat.pre <- bind_rows(relevance.dat.aa, relevance.dat.ea) %>%
  #mutate(Ancestry = factor(Ancestry, levels = c("African ancestries", "European ancestries"), labels = c("African ancestries", "European ancestries"))) %>%
  mutate(Instrument = ifelse(
    Instrument == "PGS4_DPW_GSCAN19",
    "Whole genome polygenic score",
    Instrument
  )) %>%
  mutate(Instrument = ifelse(
    Instrument == "pgs.norm",
    "Candidate SNP polygenic score",
    Instrument
  )) %>% 
  mutate(Instrument = factor(Instrument))

rsids <- levels(relevance.dat.pre$Instrument)[relevance.dat.pre$Instrument %>% levels %>% grep(pattern = "rs")]

relevance.dat <- relevance.dat.pre %>%
  mutate(Instrument = factor(Instrument, levels = c(rsids, "Candidate SNP polygenic score", "Whole genome polygenic score")))
```


```{r}
ggplot(data = relevance.dat,
       mapping = aes(
         x = Instrument,
         y = graph_F,
         color = Model
       )) +
  geom_point(alpha = geom.point.alpha) + #, size = geom.point.size) +
  geom_hline(yintercept = 10, linetype = 'dotted') +
  ggtitle("F-statistic relevance test") +
  xlab("Instrument") +
  ylab("F-statistic") +
  facet_wrap(facets = vars(Ancestry)) +
  coord_flip() +
  #theme_minimal(base_size = base.size) +
  #theme(axis.text.x = element_text(
  #  size = 5,
  #  angle = 45,
  #  hjust = 1,
  #  vjust = 1
  #))
  theme(axis.text.y = element_text(
    size = 4
  ))

#ggsave(here("results", "one_smr", paste0(Sys.Date(), "_relevance_test.png")), dpi = 300)

#ggsave(width = 1920, height = 1080, units = "px", filename = here("results", "one_smr", paste0(Sys.Date(), "_relevance_test.png")))
```


### Log-transformed drinkpweek

Fit all relevance models for naive, ancestryPCs, demographic, and full models
```{r}
res_relevance_aa_log <- fit_relevance_models(instruments = genetics, exposure = "logdrinkpweek", tidy_model = lm_mod, data = dat.aa)
```

Display and plot results
```{r}
ggplot(data = res_relevance_aa_log,
       mapping = aes(x = graph_F, color = Model, fill = Model)) +
  geom_histogram() +
  facet_wrap(facets = vars(Model)) +
  ggtitle(ggtitle(paste0("African ancestries histogram of instrument F-statistics, n = ", dim(dat.aa)[1])))

print(res_relevance_aa_log %>% filter(Instrument == "PGS4_DPW_GSCAN19"))

print(res_relevance_aa_log %>% filter(Instrument == "pgs.norm"))

ggplot(data = dat.aa, mapping = aes(x = PGS4_DPW_GSCAN19, y = logdrinkpweek)) + geom_point() + geom_smooth(method='lm', formula= y~x) +
  ggtitle(paste0("African ancestries regress log-transformed DPW on PGS4, n = ", dim(dat.aa)[1]))
```


### Outmoded analyses
```{r, eval=FALSE, include=FALSE}
res_relevance_ea_all <- fit_relevance_models(instruments = genetics, exposure = "logdrinkpweek", tidy_model = lm_mod, data = dat.all.ea)
```

Display and plot results
```{r, eval=FALSE, include=FALSE}
print(res_relevance_ea_all %>% filter(Instrument == "PGS4_DPW_GSCAN19"))

print(res_relevance_ea_all %>% filter(Instrument == "pgs.norm"))

ggplot(data = res_relevance_ea_all,
       mapping = aes(x = graph_F, color = Model, fill = Model)) +
  geom_histogram() +
  facet_wrap(facets = vars(Model)) +
  ggtitle(ggtitle(paste0("European ancestries histogram of instrument F-statistics, n = ", dim(dat.ea)[1])))



#ggplot(data = dat.all.ea, mapping = aes(x = pgs.norm, y = logdrinkpweek)) + geom_point() + geom_smooth(method='lm', formula= y~x) +
#  stat_cor() +
#  ggtitle(paste0("European ancestries regress DPW on PGS4, n = ", dim(dat.all.ea)[1]))
```

```{r, eval=FALSE, include=FALSE}
cor.test(dat.ea$pgs.norm, dat.ea$logdrinkpweek)
#cor.test(dat.all.ea$pgs.norm, dat.all.ea$logdrinkpweek)
```


#### Hypothesis: if PGSCAN GWAS excluded never drinkers or who quit before study measurement were set to missing -> better performance in non-never drinkers

```{r}
dat.ea %>% filter(everdrank == T) %>% dim
```

```{r}
table(dat.ea$R11DRINK, dat.ea$AD12)
table(dat.ea$R11DRINK, dat.ea$AD12) %>% chisq.test()
```

```{r}
table(dat.ea$everdrank, dat.ea$AD12)
table(dat.ea$everdrank, dat.ea$AD12) %>% chisq.test()
```

```{r}
res_relevance_ea_drinkers <- fit_relevance_models(instruments = genetics, exposure = "drinkpweek", tidy_model = lm_mod, data = dat.ea %>% filter(everdrank == T))
```

Display and plot results
```{r}
ggplot(data = res_relevance_ea_drinkers,
       mapping = aes(x = graph_F, color = Model, fill = Model)) +
  geom_histogram() +
  facet_wrap(facets = vars(Model)) +
  ggtitle(ggtitle(paste0("European ancestries drinkers histogram of instrument F-statistics, n = ", dim(dat.ea)[1])))

print(res_relevance_ea_drinkers %>% filter(Instrument == "PGS4_DPW_GSCAN19"))

print(res_relevance_ea_drinkers %>% filter(Instrument == "pgs.norm"))

ggplot(data = dat.ea %>% filter(everdrank == T), mapping = aes(x = pgs.norm, y = drinkpweek)) + geom_point() + geom_smooth(method='lm', formula= y~x) +
  stat_cor() +
  ggtitle(paste0("European ancestries regress DPW on PGS4, n = ", dim(dat.ea %>% filter(everdrank == T))[1]))
```

#### Hypothesis: if PGSCAN GWAS excluded not current drinkers or who quit before study measurement were set to missing -> better performance in not current drinkers

```{r}
dat.ea %>% filter(R11DRINK == 1) %>% dim
```

```{r}
table(dat.ea$R11DRINK, dat.ea$AD12)
table(dat.ea$R11DRINK, dat.ea$AD12) %>% chisq.test()
```

```{r}
table(dat.ea$everdrank, dat.ea$AD12)
table(dat.ea$everdrank, dat.ea$AD12) %>% chisq.test()
```

```{r}
res_relevance_ea_drinkers <- fit_relevance_models(instruments = genetics, exposure = "drinkpweek", tidy_model = lm_mod, data = dat.ea %>% filter(R11DRINK == 1))
```

Display and plot results
```{r}
ggplot(data = res_relevance_ea_drinkers,
       mapping = aes(x = graph_F, color = Model, fill = Model)) +
  geom_histogram() +
  facet_wrap(facets = vars(Model)) +
  ggtitle(ggtitle(paste0("European ancestries current drinkers histogram of instrument F-statistics, n = ", dim(dat.ea)[1])))

print(res_relevance_ea_drinkers %>% filter(Instrument == "PGS4_DPW_GSCAN19"))

print(res_relevance_ea_drinkers %>% filter(Instrument == "pgs.norm"))

ggplot(data = dat.ea %>% filter(R11DRINK == 1), mapping = aes(x = PGS4_DPW_GSCAN19, y = drinkpweek)) + geom_point() + geom_smooth(method='lm', formula= y~x) +
  stat_cor() +
  ggtitle(paste0("European ancestries regress DPW on PGS4, n = ", dim(dat.ea %>% filter(R11DRINK == 1))[1]))
```

https://stats.stackexchange.com/questions/64152/are-two-pearson-correlation-coefficients-different
```{r}
#cor.diff.test(dat.ea$pgs.norm, dat.ea$logdrinkpweek, dat.all.ea$pgs.norm, dat.all.ea$logdrinkpweek)

cor.diff.test = function(x1, x2, y1, y2, method="pearson") {
  cor1 = cor.test(x1, x2, method=method)
  cor2 = cor.test(y1, y2, method=method)

  r1 = cor1$estimate
  r2 = cor2$estimate
  n1 = sum(complete.cases(x1, x2))
  n2 = sum(complete.cases(y1, y2))
  fisher = ((0.5*log((1+r1)/(1-r1)))-(0.5*log((1+r2)/(1-r2))))/((1/(n1-3))+(1/(n2-3)))^0.5

  p.value = (2*(1-pnorm(abs(fisher))))

  result= list(
    "cor1" = list(
      "estimate" = as.numeric(cor1$estimate),
      "p.value" = cor1$p.value,
      "n" = n1
    ),
    "cor2" = list(
      "estimate" = as.numeric(cor2$estimate),
      "p.value" = cor2$p.value,
      "n" = n2
    ),
    "p.value.twosided" = as.numeric(p.value),
    "p.value.onesided" = as.numeric(p.value) / 2
  )
  cat(paste(sep="",
          "cor1: r=", format(result$cor1$estimate, digits=3), ", p=", format(result$cor1$p.value, digits=3), ", n=", result$cor1$n, "\n",
          "cor2: r=", format(result$cor2$estimate, digits=3), ", p=", format(result$cor2$p.value, digits=3), ", n=", result$cor2$n, "\n",
          "diffence: p(one-sided)=", format(result$p.value.onesided, digits=3), ", p(two-sided)=", format(result$p.value.twosided, digits=3), "\n"
  ))
  return(result);
}
```

#### W/o 4 DPW outliers
Fit all relevance models for naive, ancestryPCs, demographic, and full models
```{r}
res_relevance_ea_dpw_lt_40_dpw_lt_40 <- fit_relevance_models(instruments = genetics, exposure = "drinkpweek", tidy_model = lm_mod, data = dat.ea %>% filter(drinkpweek < 40))
```

Display and plot results
```{r}
ggplot(data = res_relevance_ea_dpw_lt_40_dpw_lt_40,
       mapping = aes(x = graph_F, color = Model, fill = Model)) +
  geom_histogram() +
  facet_wrap(facets = vars(Model)) +
  ggtitle(ggtitle(paste0("<40 DPW European ancestries histogram of instrument F-statistics, n = ", dim(dat.ea %>% filter(drinkpweek < 40))[1])))

print(res_relevance_ea_dpw_lt_40_dpw_lt_40 %>% filter(Instrument == "PGS4_DPW_GSCAN19"))

print(res_relevance_ea_dpw_lt_40_dpw_lt_40 %>% filter(Instrument == "pgs.norm"))

ggplot(data = dat.ea %>% filter(drinkpweek < 40), mapping = aes(x = PGS4_DPW_GSCAN19, y = drinkpweek)) + geom_point() + geom_smooth(method='lm', formula= y~x) +
  ggtitle(paste0("European ancestries regress DPW on PGS4, n = ", dim(dat.ea %>% filter(drinkpweek < 40))[1]))
```

### Log-transformed drinks per week exposure
Fit all relevance models for naive, ancestryPCs, demographic, and full models
```{r}
res_relevance_ea_log <- fit_relevance_models(instruments = genetics, exposure = "drinkpweek", tidy_model = lm_mod, data = dat.ea)
```

Display and plot results
```{r}
ggplot(data = res_relevance_ea_log,
       mapping = aes(x = graph_F, color = Model, fill = Model)) +
  geom_histogram() +
  facet_wrap(facets = vars(Model)) +
  ggtitle(ggtitle(paste0("European ancestries histogram of instrument F-statistics, n = ", dim(dat.ea)[1])))

print(res_relevance_ea_log %>% filter(Instrument == "PGS4_DPW_GSCAN19"))

print(res_relevance_ea_log %>% filter(Instrument == "pgs.norm"))

ggplot(data = dat.ea, mapping = aes(x = PGS4_DPW_GSCAN19, y = logdrinkpweek)) + geom_point() + geom_smooth(method='lm', formula= y~x) +
  ggtitle(paste0("European ancestries regress log-transformed DPW on PGS4, n = ", dim(dat.ea)[1]))
```

# Exclusion

```{r}
tidy_parsnip_map_lr <- function(fit, map.x) {
  
  fit_res <- map2(fit, genetics, ~ tidy(.x, conf.int = TRUE, exponentiate = TRUE) %>% filter(term == .y))
  names(fit_res) <- genetics
  fit_res_df <- bind_rows(fit_res, .id = "Instrument")
  
  fit_glance <- map(fit, ~ glance(.x))
  names(fit_glance) <- genetics
  fit_glance_df <- bind_rows(fit_glance, .id = "Instrument") #%>% rename("F.stat" = statistic, "F.p.value" = p.value)
  
  fit_all <- left_join(fit_res_df, fit_glance_df, by = "Instrument")
  
  return(fit_all)
}
```

## European ancestries - Dementia
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred, 22 times
```{r}
fit <-
  map(
    .x = genetics,
    .f = ~ lr_mod %>%
      fit(formula(paste0("AD12.Dementia.outcome ~ ", .x)),
          data = dat.ea)
  )

# If you want to check out the observations that have probabilities near 0 or 1
#fit.GSCAN <- lr_mod %>%
#  fit(formula("AD12.Dementia.outcome ~ PGS4_DPW_GSCAN19"),
#      data = dat.ea)
#View(predict(fit.GSCAN, dat.ea, type = "prob"))

res_dementia_ea <- tidy_parsnip_map_lr(fit) %>%
  mutate(Ancestry = "European Ancestries") %>%
  mutate(outcome = "Dementia")
```

## European ancestries - CIND
```{r}
fit <-
  map(
    .x = genetics,
    .f = ~ lr_mod %>%
      fit(formula(paste0("AD12.CIND.outcome ~ ", .x)),
          data = dat.ea)
  )

res_cind_ea <- tidy_parsnip_map_lr(fit) %>%
  mutate(Ancestry = "European Ancestries") %>%
  mutate(outcome = "CIND")
```


## African ancestries - Dementia
```{r}
fit <-
  map(
    .x = genetics,
    .f = ~ lr_mod %>%
      fit(formula(paste0("AD12.Dementia.outcome ~ ", .x)),
          data = dat.aa)
  )

# If you want to check out the observations that have probabilities near 0 or 1
#fit.GSCAN <- lr_mod %>%
#  fit(formula("AD12.Dementia.outcome ~ PGS4_DPW_GSCAN19"),
#      data = dat.aa)
#View(predict(fit.GSCAN, dat.aa, type = "prob"))

res_dementia_aa <- tidy_parsnip_map_lr(fit) %>%
  mutate(Ancestry = "African Ancestries") %>%
  mutate(outcome = "Dementia")
```

## African ancestries - CIND
```{r}
fit <-
  map(
    .x = genetics,
    .f = ~ lr_mod %>%
      fit(formula(paste0("AD12.CIND.outcome ~ ", .x)),
          data = dat.aa)
  )

res_cind_aa <- tidy_parsnip_map_lr(fit) %>%
  mutate(Ancestry = "African Ancestries") %>%
  mutate(outcome = "CIND")
```

## Exclusion figure
```{r}
instruments.to.label <- c("pgs.norm", "PGS4_DPW_GSCAN19", "rs1229984")
exclusion.dat <- bind_rows(res_cind_ea, res_dementia_ea, res_cind_aa, res_dementia_aa)
exclusion.dat$p.value.adj <- p.adjust(exclusion.dat$p.value, method = "fdr")
exclusion.dat$label[(exclusion.dat$Instrument %in% instruments.to.label)] <- T
exclusion.dat <- exclusion.dat %>%
  mutate(Instrument = ifelse(
    Instrument == "pgs.norm",
    "Candidate SNP polygenic score",
    Instrument)) %>%
  mutate(Instrument = ifelse(
    Instrument == "PGS4_DPW_GSCAN19",
    "Whole genomice polygenic socre",
    Instrument)) %>%
  mutate(label = ifelse(
    -log10(p.value) > 1.3,
    T, label
  ))
```

```{r}
ggplot(data = exclusion.dat,
       mapping = aes(
         x = estimate,
         y = -log10(p.value),
         color = outcome
       )) +
  geom_point(alpha = geom.point.alpha) + #, size = geom.point.size) +
  #ylim(c(0, 12)) +
  geom_hline(yintercept = 1.3, linetype = 'dotted') +
  geom_text_repel(aes(
        x = estimate,
        y = -log10(p.value),
        label = ifelse(label, Instrument, "")
      ), size = 3) +
  ggtitle("Logistic regression coefficient test of exclusion") +
  #labs(caption = paste0(ancestry.hits.truncated.num, " additional ancestries hits truncated")) +
  xlab("Instrument") +
  ylab("-log10 p-value") +
  facet_wrap(facets = vars(Ancestry)) +
  #theme_minimal(base_size = base.size) +
  theme(axis.text.x = element_text(
    angle = 45,
    hjust = 1,
    vjust = 1
  ))

#ggsave(filename = here("results", "one_smr", paste0(Sys.Date(), "_exclusion_test.png")))
#ggsave(width = 1920, height = 1080, units = "px", filename = here("results", "one_smr", paste0(Sys.Date(), "_exclusion_test.png")))
```

# Exogeneity
## European ancestries
##Fit instrument -> covariate
```{r}
independent.vars <- c("NAGE", "GENDER", "DEGREE", "APOE012", "R11SMOKEV", "NMARST", "R11CESD", "R11CONDE", "R11SAYRET", "AncestryPC_1_5A", "AncestryPC_1_5B", "AncestryPC_1_5C", "AncestryPC_1_5D", "AncestryPC_1_5E")

cat.vars <- c("APOE012", "NAMRST", "R11SAYRET", "DEGREE")
bin.vars <- c("GENDER", "R11SMOKEV")
cont.vars <- c("NAGE", "R11CESD", "R11CONDE", "AncestryPC_1_5A", "AncestryPC_1_5B", "AncestryPC_1_5C", "AncestryPC_1_5D", "AncestryPC_1_5E")
```

### All potential confounders with linear model F-test 

```{r}
crossed <- crossing(genetics, independent.vars)
genetics.crossed <- crossed$genetics
independent.vars.crossed <- crossed$independent.vars
```

```{r}
# Fit the linear model across all instrument and covariate combinations
fit <-
  map2(
    .x = genetics.crossed,
    .y = independent.vars.crossed,
    .f = ~ lm_mod %>%
      fit(formula(paste0(.x, " ~ ", .y)),
          data = dat.ea)
  )

# Collect the results into a data frame
# Add FDR-adjusted p-value
fit_glance <- map(fit, ~ glance(.x))
names(fit_glance) <- paste0(genetics.crossed, "+", independent.vars.crossed)
fit_glance_df_ea <-
 bind_rows(fit_glance, .id = "Instrument") %>%
 rename("F.stat" = statistic, "F.p.value" = p.value) %>%
 mutate(F.p.value.fdr = p.adjust(F.p.value, method = "fdr")) %>%
 rowwise() %>%
 mutate(predictor = strsplit(Instrument,"+", fixed = T)[[1]][2]) %>%
 mutate(Instrument = strsplit(Instrument,"+", fixed = T)[[1]][1]) %>%
  dplyr::select(Instrument, predictor, everything()) %>%
  mutate(Instrument = ifelse(
    Instrument == "PGS4_DPW_GSCAN19",
    "Whole genome polygenic score",
    Instrument
  )) %>%
  mutate(Instrument = ifelse(
    Instrument == "pgs.norm",
    "Candidate SNP polygenic score",
    Instrument
  )) %>%
  mutate(Instrument = factor(Instrument, levels = c(rsids, "Candidate SNP polygenic score", "Whole genome polygenic score")))
```

```{r}
ancestry.hits.truncated.num.ea <- fit_glance_df_ea %>% filter(-log10(F.p.value) > 12) %>% nrow %>% as.numeric()
ggplot(data = fit_glance_df_ea, mapping = aes(x = Instrument, y = -log10(F.p.value), color = predictor)) + geom_point(alpha = 0.5) + ylim(c(0, 12)) + geom_hline(yintercept = 1.3, linetype = 'dotted') + ggtitle("Simple F-test of exogeneity", subtitle = paste0("European ancestries, n = ", nrow(dat.ea))) + labs(caption = paste0(ancestry.hits.truncated.num.ea, " additional ancestries hits truncated")) + xlab("Instrument") + ylab("-log10 p-value") + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 5))
```


```{r}
fit_glance_df_ea %>% group_by(Instrument) %>% filter(F.p.value < 0.05) %>% tally()
fit_glance_df_ea %>% group_by(Instrument) %>% filter(F.p.value.fdr < 0.05) %>% tally()
fit_glance_df_ea %>% group_by(Instrument) %>% filter(F.p.value < 0.05) %>% tally() %>% pull(n) %>% hist
fit_glance_df_ea %>% group_by(Instrument) %>% filter(F.p.value.fdr < 0.05) %>% tally() %>% pull(n) %>% hist
```

```{r}
fit_glance_df_ea %>% group_by(predictor) %>% filter(F.p.value < 0.05) %>% tally()
fit_glance_df_ea %>% group_by(predictor) %>% filter(F.p.value.fdr < 0.05) %>% tally()
fit_glance_df_ea %>% group_by(predictor) %>% filter(F.p.value < 0.05) %>% tally() %>% pull(n) %>% hist
fit_glance_df_ea %>% group_by(predictor) %>% filter(F.p.value.fdr < 0.05) %>% tally() %>% pull(n) %>% hist
```

## African ancestries 
### All potential confounders with linear model F-test 
```{r}
# Fit the linear model across all instrument and covariate combinations
fit <-
  map2(
    .x = genetics.crossed,
    .y = independent.vars.crossed,
    .f = ~ lm_mod %>%
      fit(formula(paste0(.x, " ~ ", .y)),
          data = dat.aa)
  )

# Collect the results into a data frame
# Add FDR-adjusted p-value
fit_glance <- map(fit, ~ glance(.x))
names(fit_glance) <- paste0(genetics.crossed, "+", independent.vars.crossed)
fit_glance_df_aa <-
 bind_rows(fit_glance, .id = "Instrument") %>%
 rename("F.stat" = statistic, "F.p.value" = p.value) %>%
 mutate(F.p.value.fdr = p.adjust(F.p.value, method = "fdr")) %>%
 rowwise() %>%
 mutate(predictor = strsplit(Instrument,"+", fixed = T)[[1]][2]) %>%
 mutate(Instrument = strsplit(Instrument,"+", fixed = T)[[1]][1]) %>%
  dplyr::select(Instrument, predictor, everything()) %>%
  mutate(Instrument = ifelse(
    Instrument == "PGS4_DPW_GSCAN19",
    "Whole genome polygenic score",
    Instrument
  )) %>%
  mutate(Instrument = ifelse(
    Instrument == "pgs.norm",
    "Candidate SNP polygenic score",
    Instrument
  )) %>%
  mutate(Instrument = factor(Instrument, levels = c(rsids, "Candidate SNP polygenic score", "Whole genome polygenic score")))
```

```{r}
ancestry.hits.truncated.num.aa <- fit_glance_df_aa %>% filter(-log10(F.p.value) > 12) %>% nrow %>% as.numeric()
ggplot(data = fit_glance_df_aa, mapping = aes(x = Instrument, y = -log10(F.p.value), color = predictor)) + geom_point(alpha = 0.5) + ylim(c(0, 12)) + geom_hline(yintercept = 1.3, linetype = 'dotted') + ggtitle("Simple F-test of exogeneity", subtitle = paste0("African ancestries, n = ", nrow(dat.aa))) + labs(caption = paste0(ancestry.hits.truncated.num.aa, " additional ancestries hits truncated")) + xlab("Instrument") + ylab("-log10 p-value") + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 5))
```

```{r}
fit_glance_df_aa %>% group_by(Instrument) %>% filter(F.p.value < 0.05) %>% tally()
fit_glance_df_aa %>% group_by(Instrument) %>% filter(F.p.value.fdr < 0.05) %>% tally()
fit_glance_df_aa %>% group_by(Instrument) %>% filter(F.p.value < 0.05) %>% tally() %>% pull(n) %>% hist
fit_glance_df_aa %>% group_by(Instrument) %>% filter(F.p.value.fdr < 0.05) %>% tally() %>% pull(n) %>% hist
```

```{r}
fit_glance_df_aa %>% group_by(predictor) %>% filter(F.p.value < 0.05) %>% tally()
fit_glance_df_aa %>% group_by(predictor) %>% filter(F.p.value.fdr < 0.05) %>% tally()
fit_glance_df_aa %>% group_by(predictor) %>% filter(F.p.value < 0.05) %>% tally() %>% pull(n) %>% hist
fit_glance_df_aa%>% group_by(predictor) %>% filter(F.p.value.fdr < 0.05) %>% tally() %>% pull(n) %>% hist
```

## Exogeneity figures

Separate analytic figures
```{r}
fit_glance_df_aa <- fit_glance_df_aa %>%
  mutate(predictor = factor(predictor, levels = c("AncestryPC_1_5A", "AncestryPC_1_5B", "AncestryPC_1_5C", "AncestryPC_1_5D", 
"AncestryPC_1_5E", "APOE012", "DEGREE", "GENDER", "NAGE", "NMARST", 
"R11CESD", "R11CONDE", "R11SAYRET", "R11SMOKEV"),
labels = c("AncestryPC_1_5A", "AncestryPC_1_5B", "AncestryPC_1_5C", "AncestryPC_1_5D", 
"AncestryPC_1_5E", "APOE status", "Education", "Sex", "Age", "Marital status", 
"Depressive symptoms", "Comorbidities", "Retirement status", "Ever smoker"))) %>%
  rename(Predictor = predictor)

exogeneity.manhattan.aa <- ggplot(data = fit_glance_df_aa, mapping = aes(x = Instrument, y = -log10(F.p.value), color = Predictor)) + geom_point(alpha = geom.point.alpha) + ylim(c(0, 12)) + geom_hline(yintercept = 1.3, linetype = 'dotted') + ggtitle("Simple F-test of exogeneity", subtitle = paste0("African ancestries, n = ", nrow(dat.aa))) + labs(caption = paste0(ancestry.hits.truncated.num.aa, " additional ancestries hits truncated")) + xlab("Instrument") + ylab("-log10 p-value") + theme(axis.text.y = element_text(
    size = 6
  )) +
  coord_flip()
exogeneity.manhattan.aa

#ggsave(dpi = 300, filename = here("results", "one_smr", paste0(Sys.Date(), "_exogeneity_test_aa.png")))
```

Separate analytic figures
```{r}
fit_glance_df_ea <- fit_glance_df_ea %>%
  mutate(predictor = factor(predictor, levels = c("AncestryPC_1_5A", "AncestryPC_1_5B", "AncestryPC_1_5C", "AncestryPC_1_5D", 
"AncestryPC_1_5E", "APOE012", "DEGREE", "GENDER", "NAGE", "NMARST", 
"R11CESD", "R11CONDE", "R11SAYRET", "R11SMOKEV"),
labels = c("AncestryPC_1_5A", "AncestryPC_1_5B", "AncestryPC_1_5C", "AncestryPC_1_5D", 
"AncestryPC_1_5E", "APOE status", "Education", "Sex", "Age", "Marital status", 
"Depressive symptoms", "Comorbidities", "Retirement status", "Ever smoker"))) %>%
  rename(Predictor = predictor)

exogeneity.manhattan.ea <- ggplot(data = fit_glance_df_ea, mapping = aes(x = Instrument, y = -log10(F.p.value), color = Predictor)) + geom_point(alpha = geom.point.alpha) + ylim(c(0, 12)) + geom_hline(yintercept = 1.3, linetype = 'dotted') + ggtitle("Simple F-test of exogeneity", subtitle = paste0("European ancestries, n = ", nrow(dat.ea))) + labs(caption = paste0(ancestry.hits.truncated.num.ea, " additional ancestries hits truncated")) + xlab("Instrument") + ylab("-log10 p-value") + theme(axis.text.y = element_text(
    size = 6
  )) +
  coord_flip()
exogeneity.manhattan.ea

#ggsave(dpi = 300, filename = here("results", "one_smr", paste0(Sys.Date(), "_exogeneity_test_ea.png")))
```

```{r, eval = F}
base.size = 24
geom.point.size = 6
geom.point.alpha = 0.75
library(ggpubr)
#exogeneity.manhattan.ea
#exogeneity.manhattan.aa
exogeneity.manhattan.aa <- ggplot(data = fit_glance_df_aa, mapping = aes(x = Instrument, y = -log10(F.p.value), color = predictor)) + geom_point(alpha = geom.point.alpha, size = geom.point.size) + ylim(c(0, 12)) + geom_hline(yintercept = 1.3, linetype = 'dotted') + ggtitle("Simple F-test of exogeneity", subtitle = paste0("African ancestries, n = ", nrow(dat.aa))) + labs(caption = paste0(ancestry.hits.truncated.num.aa, " additional ancestries hits truncated")) + xlab("Instrument") + ylab("-log10 p-value") + theme_minimal(base_size = base.size) + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

exogeneity.manhattan.ea <- ggplot(data = fit_glance_df_ea, mapping = aes(x = Instrument, y = -log10(F.p.value), color = predictor)) + geom_point(alpha = geom.point.alpha, size = geom.point.size) + ylim(c(0, 12)) + geom_hline(yintercept = 1.3, linetype = 'dotted') + ggtitle("Simple F-test of exogeneity", subtitle = paste0("European ancestries, n = ", nrow(dat.ea))) + labs(caption = paste0(ancestry.hits.truncated.num.ea, " additional ancestries hits truncated")) + xlab("Instrument") + ylab("-log10 p-value") + theme_minimal(base_size = base.size) + theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))

#panel <- ggarrange(exogeneity.manhattan.aa, exogeneity.manhattan.ea, labels = "auto", common.legend = TRUE, legend = "bottom", font.label = list(size = 24, face = "bold", color = "black"))
#ggexport(panel, width = 1920, height = 1080, filename = here("results", "one_smr", paste0(Sys.Date(), "_exogeneity_test.png")))
```

Try facet_wrap instead
```{r}
glance_res_all <- rbind(fit_glance_df_aa %>% mutate(Ancestry = "African ancestries"), fit_glance_df_ea %>% mutate(Ancestry = "European ancestries"))
```

```{r}
ggplot(data = glance_res_all,
       mapping = aes(
         x = Instrument,
         y = -log10(F.p.value),
         color = predictor
       )) +
  geom_point(alpha = geom.point.alpha) + #, size = geom.point.size) +
  geom_hline(yintercept = -log10(0.05), linetype = 'dotted') +
  ggtitle("Simple F-test of exogeneity") +
  xlab("Instrument") +
  ylab("-log10(p-value)") +
  facet_wrap(facets = vars(Ancestry)) +
  coord_flip() +
  #theme_minimal(base_size = base.size) +
  #theme(axis.text.x = element_text(
  #  size = 5,
  #  angle = 45,
  #  hjust = 1,
  #  vjust = 1
  #))
  theme(axis.text.y = element_text(
    size = 4
  )) +
  ylim(c(0, 12))
ggsave(dpi=300, filename = here("results", "one_smr", paste0(Sys.Date(), "_exogeneity_test.png")))
```
